---
title: "Cinderhaven · Monday Morning Dashboard"
subtitle: "Velocity · Retailer P&L · Triage Tracker"
format:
  html:
    toc: false
    theme: cosmo
    embed-resources: false
    css:
      - assets/report.css
      - assets/dashboard.css
    fig-align: center
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    out.width: "100%"
---

```{r setup}
#| include: false
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(stringr); library(forcats)
  library(lubridate); library(ggplot2); library(scales)
  library(reactable); library(htmltools)
})

source("../R/00_theme.R")  # cinderhaven_palette + theme_cinderhaven_*

PROC_DIR <- file.path("..", "output", "frames")
read_p <- function(name) readRDS(file.path(PROC_DIR, paste0(name, ".rds")))

sku_master_full <- read_p("sku_master_full")
retailer_pnl    <- read_p("retailer_pnl")
raw             <- read_p("raw_tables")

# ---- shared helpers -----------------------------------------------------

dollar_short <- function(x) {
  ifelse(is.na(x), "—",
   ifelse(abs(x) >= 1e6, sprintf("$%.2fM", x / 1e6),
    ifelse(abs(x) >= 1e3, sprintf("$%.0fk", x / 1e3),
                          paste0("$", formatC(round(x), big.mark = ",", format = "d")))))
}

dollar_cell <- function(value) {
  if (is.na(value)) return("—")
  if (abs(value) >= 1e6) sprintf("$%.2fM", value / 1e6)
  else if (abs(value) >= 1e3) sprintf("$%.0fk", value / 1e3)
  else paste0("$", formatC(round(value), big.mark = ",", format = "d"))
}

ramp_red <- function(x) {
  if (is.na(x)) return("#FFFFFF")
  grDevices::colorRampPalette(c("#FFFFFF", cinderhaven_palette$red))(101)[
    pmin(101, pmax(1, round(x * 100) + 1))]
}
ramp_teal <- function(x) {
  if (is.na(x)) return("#FFFFFF")
  grDevices::colorRampPalette(c("#FFFFFF", cinderhaven_palette$teal))(101)[
    pmin(101, pmax(1, round(x * 100) + 1))]
}

# ---- velocity (SKU × retailer, 4-week vs 12-week vs prior 4-week) -------
# Mirrors the canonical computation in R/06_excel_workbook.R so the
# dashboard and the Excel workbook always agree.

scan <- raw$scan_data |>
  mutate(week_ending = ymd(week_ending)) |>
  left_join(raw$stores |> select(store_id, retailer), by = "store_id") |>
  filter(!is.na(retailer))

last_week   <- max(scan$week_ending, na.rm = TRUE)
cut_4w      <- last_week - weeks(4)
cut_12w     <- last_week - weeks(12)
cut_8w_back <- last_week - weeks(8)

agg_window <- function(df) df |>
  group_by(sku, retailer) |>
  summarise(units  = sum(units_sold),
            stores = n_distinct(store_id),
            weeks  = n_distinct(week_ending),
            .groups = "drop")

v_4w   <- agg_window(filter(scan, week_ending >  cut_4w)) |>
  rename(units_4w = units, stores_4w = stores, weeks_4w = weeks)
v_12w  <- agg_window(filter(scan, week_ending >  cut_12w)) |>
  rename(units_12w = units, stores_12w = stores, weeks_12w = weeks)
v_prev <- agg_window(filter(scan, week_ending <= cut_4w &
                                   week_ending >  cut_8w_back)) |>
  rename(units_prev4 = units, stores_prev4 = stores, weeks_prev4 = weeks)

velocity <- v_4w |>
  full_join(v_12w,  by = c("sku", "retailer")) |>
  full_join(v_prev, by = c("sku", "retailer")) |>
  left_join(sku_master_full |>
              select(sku, product_name, product_line,
                     issue_count, data_quality_score),
            by = "sku") |>
  mutate(
    ups_4w   = units_4w   / pmax(stores_4w   * weeks_4w,   1),
    ups_12w  = units_12w  / pmax(stores_12w  * weeks_12w,  1),
    ups_prev = units_prev4 / pmax(stores_prev4 * weeks_prev4, 1),
    pct_change = ifelse(is.na(ups_prev) | ups_prev == 0, NA_real_,
                        ups_4w / ups_prev - 1),
    quality_tier = case_when(
      data_quality_score >= 87.5 ~ "Best 25%",
      data_quality_score >= 75   ~ "Above average",
      data_quality_score >= 62.5 ~ "Below average",
      TRUE                       ~ "Worst 25%"),
    quality_tier = factor(quality_tier,
      levels = c("Worst 25%", "Below average", "Above average", "Best 25%")),
    dq_flag = ifelse(issue_count >= 3, "REVIEW", "OK")
  ) |>
  filter(retailer %in% c("Walmart", "Costco", "UNFI", "Whole Foods")) |>
  arrange(desc(units_12w))

# est_fix_hours, savings_per_hour, and still_broken are read directly
# from sku_master_full (built in R/02_build_frames.R). The dashboard does
# not recompute them.
```

::: {.panel-tabset}

## Tab 1: Velocity Report

> **What this replaces:** The Monday morning Excel workbook built from four
> separate retailer CSV exports, reconciled by hand, with store count
> definitions that don't always agree with the broker's numbers. This table
> pulls from the same scan data, applies one consistent store count
> definition, and updates when the underlying data refreshes.
>
> **How to read it:** Each row is one SKU at one retailer. The 4-week and
> 12-week velocity columns show units per store per week. The % change
> column compares the two. A negative number means the SKU is decelerating.
> The data quality flag marks SKUs with active defects in the product
> master. Filter by retailer, product line, or quality tier using the
> controls above the table.
>
> **What to look for:** SKUs where velocity is dropping and the data
> quality flag is red. A velocity decline at a clean SKU is a commercial
> problem. A velocity decline at a flagged SKU may be a data problem
> disguised as a commercial one.

```{r velocity-table}
vel_tbl <- velocity |>
  transmute(
    SKU            = sku,
    Product        = product_name,
    Retailer       = retailer,
    Line           = product_line,
    `Quality tier` = quality_tier,
    Stores         = coalesce(stores_4w, stores_12w, 0L),
    `4w u/store/wk`  = round(ups_4w, 2),
    `12w u/store/wk` = round(ups_12w, 2),
    `% change`     = pct_change,
    `DQ flag`      = dq_flag
  )

retailer_filter <- function(values, filterValue) {
  if (is.null(filterValue) || filterValue == "All") TRUE
  else values == filterValue
}

reactable(
  vel_tbl,
  searchable      = TRUE,
  filterable      = TRUE,
  highlight       = TRUE,
  striped         = TRUE,
  fullWidth       = FALSE,
  bordered        = TRUE,
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100, 500),
  defaultSorted   = list(`12w u/store/wk` = "desc"),
  defaultColDef   = colDef(
    headerStyle = list(background = cinderhaven_palette$bg_pale,
                       fontWeight = 600,
                       color      = cinderhaven_palette$text)
  ),
  columns = list(
    SKU      = colDef(width = 95),
    Product  = colDef(width = 220),
    Retailer = colDef(width = 110, filterInput = function(values, name) {
      tags$select(
        onchange = sprintf(
          "Reactable.setFilter('velocity-table', '%s', event.target.value || undefined)",
          name),
        tags$option(value = "", "All"),
        lapply(sort(unique(values)), tags$option),
        style = "width:100%; height:24px"
      )
    }),
    Line     = colDef(width = 150, filterInput = function(values, name) {
      tags$select(
        onchange = sprintf(
          "Reactable.setFilter('velocity-table', '%s', event.target.value || undefined)",
          name),
        tags$option(value = "", "All"),
        lapply(sort(unique(values)), tags$option),
        style = "width:100%; height:24px"
      )
    }),
    `Quality tier` = colDef(
      width = 130,
      filterInput = function(values, name) {
        tags$select(
          onchange = sprintf(
            "Reactable.setFilter('velocity-table', '%s', event.target.value || undefined)",
            name),
          tags$option(value = "", "All"),
          lapply(levels(values), tags$option),
          style = "width:100%; height:24px"
        )
      },
      style = function(value) {
        v <- as.character(value)
        col <- switch(v,
          "Worst 25%"     = cinderhaven_palette$red,
          "Below average" = cinderhaven_palette$coral,
          "Above average" = cinderhaven_palette$blue_muted,
          "Best 25%"      = cinderhaven_palette$teal,
          NULL)
        if (is.null(col)) NULL else list(color = col, fontWeight = "600")
      }
    ),
    Stores   = colDef(width = 80, align = "right"),
    `4w u/store/wk`  = colDef(width = 120, align = "right",
                              format = colFormat(digits = 2)),
    `12w u/store/wk` = colDef(width = 120, align = "right",
                              format = colFormat(digits = 2)),
    `% change` = colDef(
      width = 100, align = "right",
      cell = function(value) {
        if (is.na(value)) return("—")
        sprintf("%+.0f%%", value * 100)
      },
      style = function(value) {
        if (is.na(value)) return(NULL)
        if (value < -0.05)
          list(background = ramp_red(pmin(1, abs(value) * 2)),
               fontWeight = "600")
        else if (value > 0.05)
          list(background = ramp_teal(pmin(1, value * 2)))
        else NULL
      }
    ),
    `DQ flag` = colDef(
      width = 90, align = "center",
      style = function(value) {
        if (is.na(value)) return(NULL)
        if (value == "REVIEW")
          list(background = cinderhaven_palette$red, color = "white",
               fontWeight = "600")
        else
          list(background = cinderhaven_palette$teal, color = "white",
               fontWeight = "600")
      }
    )
  ),
  elementId = "velocity-table",
  theme = reactableTheme(
    borderColor    = "#d8d8d8",
    stripedColor   = "#fafafa",
    highlightColor = "#f3eaea",
    cellPadding    = "6px 10px",
    style          = list(
      fontFamily = "'Helvetica Neue', Inter, system-ui, sans-serif",
      fontSize   = "0.88rem")
  )
)
```

## Tab 2: Retailer P&L

> **What this shows:** Net contribution by retailer after trade spend and
> chargebacks. The waterfall chart breaks gross revenue into the components
> that reduce it. The table below has the numbers.
>
> **How to read it:** Gross revenue is the top bar. Trade spend and
> chargebacks subtract from it. Net contribution is what remains. Filter by
> time period to compare quarters.
>
> **What to look for:** The gap between gross revenue rank and net margin
> rank. Walmart is #1 in gross dollars and #4 in net margin percentage.
> Whole Foods is #4 in gross dollars and #1 in net margin. Chargebacks are
> the only margin component in this chart that is entirely within
> Cinderhaven's operational control.

```{r retailer-waterfall}
#| fig-width: 11
#| fig-height: 5

contracted <- c("Walmart", "UNFI", "Whole Foods", "Costco")

# Inflate the small segments just enough to keep the dollar labels
# readable on Costco / UNFI / Whole Foods. Labels still print real $.
MIN_TRADE_VISUAL <- 0.7e6
MIN_CB_VISUAL    <- 0.5e6

pnl <- retailer_pnl |>
  filter(retailer %in% contracted) |>
  arrange(desc(gross_revenue)) |>
  mutate(retailer     = factor(retailer, levels = retailer),
         trade_visual = pmax(trade_spend_total, MIN_TRADE_VISUAL),
         cb_visual    = pmax(chargeback_total,  MIN_CB_VISUAL),
         bar_end      = net_contribution + trade_visual + cb_visual)

pnl_long <- pnl |>
  transmute(retailer,
            "Net contribution" = net_contribution,
            "Trade spend"      = trade_visual,
            "Chargebacks"      = cb_visual) |>
  pivot_longer(-retailer, names_to = "component", values_to = "value") |>
  mutate(component = factor(component,
    levels = c("Net contribution", "Trade spend", "Chargebacks")))

stack_palette <- c(
  "Net contribution" = "#D0D0D0",
  "Trade spend"      = "#888888",
  "Chargebacks"      = cinderhaven_palette$red)

lab_net   <- pnl |> transmute(retailer,
              x = net_contribution / 2,
              label = dollar_short(net_contribution))
lab_trade <- pnl |> transmute(retailer,
              x = net_contribution + trade_visual / 2,
              label = dollar_short(trade_spend_total))
lab_cb    <- pnl |> transmute(retailer,
              x = net_contribution + trade_visual + cb_visual / 2,
              label = dollar_short(chargeback_total))
lab_pct   <- pnl |> transmute(retailer, bar_end,
              label = sprintf("%.0f%% net margin",
                              100 * net_margin_pct_of_gross_revenue))

ggplot(pnl_long, aes(value, fct_rev(retailer), fill = component)) +
  geom_col(width = 0.65, color = "white", linewidth = 0.5,
           position = position_stack(reverse = TRUE)) +
  geom_text(data = lab_net,   aes(x = x, y = fct_rev(retailer), label = label),
            inherit.aes = FALSE, size = 3.0,
            color = cinderhaven_palette$text, fontface = "bold") +
  geom_text(data = lab_trade, aes(x = x, y = fct_rev(retailer), label = label),
            inherit.aes = FALSE, size = 3.0, color = "white", fontface = "bold") +
  geom_text(data = lab_cb,    aes(x = x, y = fct_rev(retailer), label = label),
            inherit.aes = FALSE, size = 3.0, color = "white", fontface = "bold") +
  geom_text(data = lab_pct, aes(x = bar_end, y = fct_rev(retailer), label = label),
            inherit.aes = FALSE, hjust = -0.06, size = 3.4, fontface = "bold",
            color = cinderhaven_palette$text) +
  scale_fill_manual(values = stack_palette, name = NULL,
                    breaks = c("Net contribution", "Trade spend", "Chargebacks")) +
  scale_x_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"),
                     expand = expansion(mult = c(0, 0.20))) +
  labs(x = NULL, y = NULL,
       caption = src_caption("Cinderhaven sales + cost data + chargeback ledger, trailing 12 months")) +
  theme_cinderhaven_horizontal() +
  theme(axis.text.y = element_text(size = 11, face = "bold"))
```

```{r retailer-pnl-table}
pnl_tbl <- retailer_pnl |>
  filter(retailer %in% contracted) |>
  arrange(desc(gross_revenue)) |>
  transmute(
    Retailer            = retailer,
    `Gross revenue`     = gross_revenue,
    `Trade spend`       = trade_spend_total,
    `Trade % of rev`    = trade_spend_pct_of_revenue,
    `Chargebacks`       = chargeback_total,
    `Cb % of rev`       = chargeback_pct_of_revenue,
    `Net contribution`  = net_contribution,
    `Net margin %`      = net_margin_pct_of_gross_revenue
  )

reactable(
  pnl_tbl,
  searchable = FALSE, filterable = FALSE, sortable = TRUE,
  fullWidth = FALSE, bordered = TRUE, striped = TRUE, highlight = TRUE,
  defaultPageSize = 4,
  defaultColDef = colDef(
    headerStyle = list(background = cinderhaven_palette$bg_pale,
                       fontWeight = 600,
                       color      = cinderhaven_palette$text)
  ),
  columns = list(
    Retailer = colDef(width = 130, style = list(fontWeight = "600")),
    `Gross revenue`    = colDef(align = "right",
                                cell = function(v) dollar_cell(v)),
    `Trade spend`      = colDef(align = "right",
                                cell = function(v) dollar_cell(v)),
    `Trade % of rev`   = colDef(align = "right",
                                cell = function(v) sprintf("%.1f%%", v * 100)),
    `Chargebacks`      = colDef(align = "right",
                                cell = function(v) dollar_cell(v),
                                style = function(v) {
                                  list(background = ramp_red(
                                    v / max(pnl_tbl$Chargebacks, na.rm = TRUE)))
                                }),
    `Cb % of rev`      = colDef(align = "right",
                                cell = function(v) sprintf("%.2f%%", v * 100)),
    `Net contribution` = colDef(align = "right",
                                cell = function(v) dollar_cell(v)),
    `Net margin %`     = colDef(align = "right",
                                cell = function(v) sprintf("%.1f%%", v * 100),
                                style = function(v) {
                                  list(background = ramp_teal(
                                    (v - 0.75) / 0.15),
                                    fontWeight  = "600")
                                })
  ),
  theme = reactableTheme(
    borderColor    = "#d8d8d8",
    stripedColor   = "#fafafa",
    cellPadding    = "8px 12px",
    style          = list(
      fontFamily = "'Helvetica Neue', Inter, system-ui, sans-serif",
      fontSize   = "0.92rem")
  )
)
```

## Tab 3: Triage Tracker

> **What this shows:** All 90 SKUs ranked by fix priority. The composite
> score weights revenue (40%), data quality (30%), and chargeback exposure
> (30%). Effort is shown separately.
>
> **How to read it:** Sort by composite to see what matters most to the
> business. Sort by estimated fix hours to see what's fastest. Sort by
> savings per hour to see the highest-return fixes. The "still broken"
> column shows the specific defect generating active chargebacks. If that
> column is blank, the SKU has no active chargeback-linked defects.
>
> **What to do with it:** Use the composite column to set the quarterly
> cleanup agenda. Use the effort column to fill a Tuesday afternoon. Start
> with any SKU where savings per hour exceeds $100. There are nine of them.
> They take 90 minutes combined.

```{r triage-tracker}
# Triage columns (composite priority, fix hours, savings/hour, still-broken)
# come straight from sku_master_full — see R/02_build_frames.R.
triage <- sku_master_full |>
  arrange(desc(fix_priority_score)) |>
  transmute(
    SKU                   = sku,
    Product               = product_name,
    `TTM revenue`         = ttm_revenue,
    `Quality score`       = round(data_quality_score, 1),
    `Chargebacks 18mo`    = chargeback_total,
    `Composite priority`  = fix_priority_score,
    `Est. fix hours`      = round(est_fix_hours, 2),
    `$ saved / hour`      = savings_per_hour,
    `Still broken`        = still_broken
  )

max_cb     <- max(triage$`Chargebacks 18mo`, na.rm = TRUE)
max_priori <- 100
max_sph    <- max(triage$`$ saved / hour`,    na.rm = TRUE)

reactable(
  triage,
  searchable      = TRUE,
  filterable      = TRUE,
  highlight       = TRUE,
  striped         = TRUE,
  fullWidth       = FALSE,
  bordered        = TRUE,
  defaultPageSize = 18,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(18, 30, 90),
  defaultSorted   = list(`Composite priority` = "desc"),
  defaultColDef   = colDef(
    headerStyle = list(background = cinderhaven_palette$bg_pale,
                       fontWeight = 600,
                       color      = cinderhaven_palette$text)
  ),
  columns = list(
    SKU     = colDef(width = 90),
    Product = colDef(width = 280),
    `TTM revenue` = colDef(
      align = "right",
      cell  = function(v) dollar_cell(v),
      style = function(v) {
        list(background = ramp_teal(v / max(triage$`TTM revenue`, na.rm = TRUE)))
      }
    ),
    `Quality score` = colDef(
      align = "right",
      style = function(v) list(background = ramp_red(1 - v / 100))
    ),
    `Chargebacks 18mo` = colDef(
      align = "right",
      cell  = function(v) dollar_cell(v),
      style = function(v) list(background = ramp_red(v / max_cb))
    ),
    `Composite priority` = colDef(
      align = "right",
      style = function(v) list(background = ramp_red(v / max_priori),
                                fontWeight  = "600")
    ),
    `Est. fix hours` = colDef(
      align = "right",
      cell  = function(v) {
        if (is.na(v) || v == 0) "—" else sprintf("%.2f", v)
      }
    ),
    `$ saved / hour` = colDef(
      align = "right",
      cell  = function(v) {
        if (is.na(v)) return("—")
        if (v >= 1e3) sprintf("$%.1fk/hr", v / 1e3)
        else paste0("$", formatC(round(v), big.mark = ","), "/hr")
      },
      style = function(v) {
        if (is.na(v)) list(background = "#FFFFFF")
        else list(background = ramp_teal(v / max_sph))
      }
    ),
    `Still broken` = colDef(
      width = 310,
      cell = function(v) if (is.na(v)) "" else v,
      style = function(v) {
        if (is.na(v)) NULL
        else list(color = cinderhaven_palette$red, fontWeight = "500")
      }
    )
  ),
  theme = reactableTheme(
    borderColor    = "#d8d8d8",
    stripedColor   = "#fafafa",
    highlightColor = "#f3eaea",
    cellPadding    = "6px 10px",
    style          = list(
      fontFamily = "'Helvetica Neue', Inter, system-ui, sans-serif",
      fontSize   = "0.88rem")
  )
)
```

:::
