---
title: "Cinderhaven · Monday Morning Dashboard"
subtitle: "Velocity · Retailer P&L · Triage Tracker"
format:
  html:
    toc: false
    theme: [cosmo, assets/lailara.scss]
    embed-resources: false
    css:
      - assets/report.css
      - assets/dashboard.css
    fig-align: center
    page-layout: full
execute:
  echo: false
  warning: false
  message: false
knitr:
  opts_chunk:
    out.width: "100%"
---

```{r setup}
#| include: false
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(stringr); library(forcats)
  library(lubridate); library(ggplot2); library(scales)
  library(reactable); library(htmltools)
})

source("../R/00_theme.R")  # cinderhaven_palette + theme_cinderhaven_*

PROC_DIR <- file.path("..", "output", "frames")
read_p <- function(name) readRDS(file.path(PROC_DIR, paste0(name, ".rds")))

sku_master_full <- read_p("sku_master_full")
retailer_pnl    <- read_p("retailer_pnl")

# ---- shared helpers -----------------------------------------------------

dollar_short <- fmt_dollar_short

dollar_cell <- function(value) {
  if (is.na(value)) return("—")
  if (abs(value) >= 1e6) sprintf("$%.2fM", value / 1e6)
  else if (abs(value) >= 1e3) sprintf("$%.0fk", value / 1e3)
  else paste0("$", formatC(round(value), big.mark = ",", format = "d"))
}


# ---- velocity (canonical frame from R/02_build_frames.R) ----------------

velocity <- read_p("velocity") |>
  rename(ups_4w = ups_per_wk_4wk, ups_12w = ups_per_wk_12wk,
         ups_prev = ups_per_w_prev4, pct_change = ups_pct_change_4w_vs_prev) |>
  mutate(
    quality_tier = case_when(
      data_quality_score >= 87.5 ~ "Best 25%",
      data_quality_score >= 75   ~ "Above average",
      data_quality_score >= 62.5 ~ "Below average",
      TRUE                       ~ "Worst 25%"),
    quality_tier = factor(quality_tier,
      levels = c("Worst 25%", "Below average", "Above average", "Best 25%")),
    dq_flag = ifelse(issue_count >= 3, "REVIEW", "OK")
  ) |>
  filter(retailer %in% c("Walmart", "Costco", "Whole Foods")) |>
  arrange(desc(dollars_12w))
```

::: {.panel-tabset}

## Tab 1: Velocity Report

> **What this replaces:** The Monday morning Excel workbook built from four
> separate retailer CSV exports, reconciled by hand, with store count
> definitions that don't always agree with the broker's numbers. This table
> pulls from the same scan data, applies one consistent store count
> definition, and updates when the underlying data refreshes.
>
> **How to read it:** Each row is one SKU at one retailer. The 4-week and
> 12-week velocity columns show units per store per week. The % change
> column compares the two. A negative number means the SKU is decelerating.
> The data quality flag marks SKUs with active defects in the product
> master. Filter by retailer, product line, or quality tier using the
> controls above the table.
>
> **What to look for:** SKUs where velocity is dropping and the data
> quality flag is red. A velocity decline at a clean SKU is a commercial
> problem. A velocity decline at a flagged SKU may be a data problem
> disguised as a commercial one.

```{r velocity-table}
vel_tbl <- velocity |>
  transmute(
    SKU            = sku,
    Product        = product_name,
    Retailer       = retailer,
    Line           = product_line,
    `Quality tier` = quality_tier,
    Stores         = coalesce(stores_4w, stores_12w, 0L),
    `4w u/store/wk`  = round(ups_4w, 2),
    `12w u/store/wk` = round(ups_12w, 2),
    `% change`     = pct_change,
    `DQ flag`      = dq_flag
  )

retailer_filter <- function(values, filterValue) {
  if (is.null(filterValue) || filterValue == "All") TRUE
  else values == filterValue
}

reactable(
  vel_tbl,
  searchable      = TRUE,
  filterable      = TRUE,
  highlight       = TRUE,
  striped         = TRUE,
  fullWidth       = TRUE,
  bordered        = TRUE,
  defaultPageSize = 20,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(20, 50, 100, 500),
  defaultSorted   = list(`12w u/store/wk` = "desc"),
  defaultColDef   = colDef(
    headerStyle = list(background = "#1f2e7a",
                       fontWeight = 600,
                       color      = "#ffffff")
  ),
  columns = list(
    SKU      = colDef(minWidth = 100, class = "col-sku", headerClass = "col-sku"),
    Product  = colDef(minWidth = 220),
    Retailer = colDef(minWidth = 110, filterInput = function(values, name) {
      tags$select(
        onchange = sprintf(
          "Reactable.setFilter('velocity-table', '%s', event.target.value || undefined)",
          name),
        tags$option(value = "", "All"),
        lapply(sort(unique(values)), tags$option),
        style = "width:100%; height:24px"
      )
    }),
    Line     = colDef(minWidth = 135, filterInput = function(values, name) {
      tags$select(
        onchange = sprintf(
          "Reactable.setFilter('velocity-table', '%s', event.target.value || undefined)",
          name),
        tags$option(value = "", "All"),
        lapply(sort(unique(values)), tags$option),
        style = "width:100%; height:24px"
      )
    }),
    `Quality tier` = colDef(
      minWidth = 130,
      filterInput = function(values, name) {
        tags$select(
          onchange = sprintf(
            "Reactable.setFilter('velocity-table', '%s', event.target.value || undefined)",
            name),
          tags$option(value = "", "All"),
          lapply(levels(values), tags$option),
          style = "width:100%; height:24px"
        )
      }
    ),
    Stores   = colDef(minWidth = 80, align = "right"),
    `4w u/store/wk`  = colDef(minWidth = 150, align = "right",
                              format = colFormat(digits = 2)),
    `12w u/store/wk` = colDef(minWidth = 140, align = "right",
                              format = colFormat(digits = 2)),
    `% change` = colDef(
      minWidth = 110, align = "right",
      cell = function(value) {
        if (is.na(value)) return("—")
        sprintf("%+.0f%%", value * 100)
      },
      style = function(value) {
        if (is.na(value)) return(NULL)
        if (value < -0.05) list(fontWeight = "600")
        else NULL
      }
    ),
    `DQ flag` = colDef(
      minWidth = 95, align = "center",
      style = function(value) {
        if (is.na(value)) return(NULL)
        if (value == "REVIEW")
          list(background = "#b82d4a", color = "white",
               fontWeight = "600")
        else
          list(background = "#158f75", color = "white",
               fontWeight = "600")
      }
    )
  ),
  elementId = "velocity-table",
  theme = reactableTheme(
    borderColor    = "#d9d9d9",
    stripedColor   = "#f2f2f2",
    highlightColor = "#e8eaf4",
    cellPadding    = "3px 5px",
    cellStyle      = list(fontSize = "11px", padding = "3px 5px"),
    headerStyle    = list(fontSize = "11px", padding = "3px 5px"),
    style          = list(
      fontFamily = "'Helvetica Neue', Inter, system-ui, sans-serif",
      fontSize   = "11px")
  )
)
```

## Tab 2: Retailer P&L

> **What this shows:** Gross revenue, trade spend, chargebacks, and net
> contribution by retailer — four separate bars per retailer so every
> number is readable.
>
> **How to read it:** Gross revenue and net contribution are the bookends.
> Trade spend and chargebacks are the two things subtracted in between.
>
> **What to look for:** The gap between gross revenue rank and net margin
> rank. Walmart is #1 in gross dollars and #4 in net margin percentage.
> Whole Foods is #4 in gross dollars and #1 in net margin. Chargebacks are
> the only margin component in this chart that is entirely within
> Cinderhaven's operational control.

```{r retailer-waterfall}
#| fig-width: 12
#| fig-height: 7
#| fig-format: svg

contracted <- c("Walmart", "Whole Foods", "Costco")

pnl <- retailer_pnl |>
  filter(retailer %in% contracted) |>
  arrange(desc(gross_revenue)) |>
  mutate(retailer = factor(retailer, levels = retailer))

pnl_long <- pnl |>
  transmute(retailer,
            "Gross revenue"    = gross_revenue,
            "Trade spend"      = trade_spend_total,
            "Chargebacks"      = chargeback_total,
            "Net contribution" = net_contribution) |>
  pivot_longer(-retailer, names_to = "component", values_to = "value") |>
  mutate(component = factor(component,
    levels = c("Gross revenue", "Trade spend", "Chargebacks", "Net contribution")))

pnl_long$bar_fill <- ifelse(pnl_long$component == "Chargebacks", "#b82d4a", "#1f2e7a")
pnl_long$label <- dollar_short(pnl_long$value)

ggplot(pnl_long, aes(x = value, y = fct_rev(retailer), fill = bar_fill)) +
  geom_col(width = 0.55) +
  geom_text(aes(label = label), color = "white",
            hjust = 1.1, fontface = "bold", size = 3.6) +
  scale_fill_identity() +
  scale_x_continuous(labels = label_dollar(scale = 1e-6, suffix = "M"),
                     expand = expansion(mult = c(0, 0.08))) +
  facet_wrap(~ component, ncol = 1, scales = "free_x") +
  labs(x = NULL, y = NULL,
       caption = src_caption("Cinderhaven sales + cost data + chargeback ledger, trailing 12 months")) +
  theme_cinderhaven_horizontal() +
  theme(axis.text.y = element_text(size = 11),
        strip.text  = element_text(size = 12, face = "bold",
                                   color = "#0d0d0d",
                                   family = "Playfair Display"))
```

```{r retailer-pnl-table}
pnl_tbl <- retailer_pnl |>
  filter(retailer %in% contracted) |>
  arrange(desc(gross_revenue)) |>
  transmute(
    Retailer            = retailer,
    `Gross revenue`     = gross_revenue,
    `Trade spend`       = trade_spend_total,
    `Trade % of rev`    = trade_spend_pct_of_revenue,
    `Chargebacks`       = chargeback_total,
    `Cb % of rev`       = chargeback_pct_of_revenue,
    `Net contribution`  = net_contribution,
    `Net margin %`      = net_margin_pct_of_gross_revenue
  )

reactable(
  pnl_tbl,
  searchable = FALSE, filterable = FALSE, sortable = TRUE,
  fullWidth = TRUE, bordered = TRUE, striped = TRUE, highlight = TRUE,
  defaultPageSize = 4,
  defaultColDef = colDef(
    minWidth = 120,
    headerStyle = list(background = "#1f2e7a",
                       fontWeight = 600,
                       color      = "#ffffff")
  ),
  columns = list(
    Retailer = colDef(minWidth = 130, style = list(fontWeight = "600")),
    `Gross revenue`    = colDef(minWidth = 130, align = "right",
                                cell = function(v) dollar_cell(v)),
    `Trade spend`      = colDef(minWidth = 120, align = "right",
                                cell = function(v) dollar_cell(v)),
    `Trade % of rev`   = colDef(minWidth = 130, align = "right",
                                cell = function(v) sprintf("%.1f%%", v * 100)),
    `Chargebacks`      = colDef(minWidth = 120, align = "right",
                                cell = function(v) dollar_cell(v)),
    `Cb % of rev`      = colDef(minWidth = 110, align = "right",
                                cell = function(v) sprintf("%.2f%%", v * 100)),
    `Net contribution` = colDef(minWidth = 140, align = "right",
                                cell = function(v) dollar_cell(v)),
    `Net margin %`     = colDef(minWidth = 120, align = "right",
                                cell = function(v) sprintf("%.1f%%", v * 100),
                                style = function(v) list(fontWeight = "600"))
  ),
  theme = reactableTheme(
    borderColor    = "#d9d9d9",
    stripedColor   = "#f2f2f2",
    cellPadding    = "3px 5px",
    cellStyle      = list(fontSize = "11px", padding = "3px 5px"),
    headerStyle    = list(fontSize = "11px", padding = "3px 5px"),
    style          = list(
      fontFamily = "'Helvetica Neue', Inter, system-ui, sans-serif",
      fontSize   = "11px")
  )
)
```

## Tab 3: Triage Tracker

> **What this shows:** All `r nrow(sku_master_full)` SKUs ranked by fix priority. The composite
> score weights revenue (40%), data quality (30%), and chargeback exposure
> (30%). Effort is shown separately.
>
> **How to read it:** Sort by composite to see what matters most to the
> business. Sort by estimated fix hours to see what's fastest. Sort by
> savings per hour to see the highest-return fixes. The "still broken"
> column shows the specific defect generating active chargebacks. If that
> column is blank, the SKU has no active chargeback-linked defects.
>
> **What to do with it:** Use the composite column to set the quarterly
> cleanup agenda. Use the effort column to fill a Tuesday afternoon. Start
> with the six quick-win SKUs (20 minutes each). They take two hours
> combined and each recovers over $15,000 per fix-hour.

```{r triage-tracker}
# Triage columns (composite priority, fix hours, savings/hour, still-broken)
# come straight from sku_master_full — see R/02_build_frames.R.
triage <- sku_master_full |>
  arrange(desc(savings_per_hour > 0), desc(savings_per_hour)) |>
  transmute(
    SKU                   = sku,
    Product               = product_name,
    `TTM revenue`         = ttm_revenue,
    `Quality score`       = round(data_quality_score, 1),
    `Chargebacks 36mo`    = chargeback_total,
    `Composite priority`  = fix_priority_score,
    `Est. fix hours`      = round(est_fix_hours, 2),
    `$ saved / hour`      = savings_per_hour,
    `Still broken`        = still_broken
  )

max_cb     <- max(triage$`Chargebacks 36mo`, na.rm = TRUE)
max_priori <- 100
max_sph    <- max(triage$`$ saved / hour`,    na.rm = TRUE)

reactable(
  triage,
  searchable      = TRUE,
  filterable      = TRUE,
  highlight       = TRUE,
  striped         = TRUE,
  fullWidth       = TRUE,
  bordered        = TRUE,
  defaultPageSize = 18,
  showPageSizeOptions = TRUE,
  pageSizeOptions = c(18, 30, nrow(sku_master_full)),
  defaultSorted   = list(`$ saved / hour` = "desc"),
  defaultColDef   = colDef(
    headerStyle = list(background = "#1f2e7a",
                       fontWeight = 600,
                       color      = "#ffffff")
  ),
  columns = list(
    SKU     = colDef(minWidth = 95, class = "col-sku", headerClass = "col-sku"),
    Product = colDef(minWidth = 175),
    `TTM revenue` = colDef(
      minWidth = 105, align = "right",
      cell  = function(v) dollar_cell(v)
    ),
    `Quality score` = colDef(minWidth = 105, align = "right"),
    `Chargebacks 36mo` = colDef(
      minWidth = 130, align = "right",
      cell  = function(v) dollar_cell(v)
    ),
    `Composite priority` = colDef(
      minWidth = 135, align = "right",
      style = function(v) list(fontWeight = "600")
    ),
    `Est. fix hours` = colDef(
      minWidth = 105, align = "right",
      cell  = function(v) {
        if (is.na(v) || v == 0) "—" else sprintf("%.2f", v)
      }
    ),
    `$ saved / hour` = colDef(
      minWidth = 100, align = "right",
      cell  = function(v) {
        if (is.na(v)) return("—")
        if (v >= 1e3) sprintf("$%.1fk/hr", v / 1e3)
        else paste0("$", formatC(round(v), big.mark = ","), "/hr")
      }
    ),
    `Still broken` = colDef(
      minWidth = 220,
      html = TRUE,
      class = "col-still-broken",
      headerClass = "col-still-broken",
      cell = function(v) if (is.na(v)) "" else gsub("; ", ";<br>", v),
      style = function(v) {
        base <- list(whiteSpace = "normal")
        if (!is.na(v)) {
          base$color <- cinderhaven_palette$red
          base$fontWeight <- "500"
        }
        base
      }
    )
  ),
  theme = reactableTheme(
    borderColor    = "#d9d9d9",
    stripedColor   = "#f2f2f2",
    highlightColor = "#e8eaf4",
    cellPadding    = "3px 5px",
    cellStyle      = list(fontSize = "11px", padding = "3px 5px"),
    headerStyle    = list(fontSize = "11px", padding = "3px 5px"),
    style          = list(
      fontFamily = "'Helvetica Neue', Inter, system-ui, sans-serif",
      fontSize   = "11px")
  )
)
```

:::
